---
title: "HW 2"
output: github_document
---
```{r}
library(readr)
```
### 2) Import the data in R
```{r}
base <- read.csv(file = "/Users/deborahchang/Desktop/stat133/stat133-hws-fall17/hw02/data/nba2017-player-statistics.csv", header = TRUE, colClasses = c("Player" = "character", "Team"  = "character", "Position" = "factor", "Experience" ="character", "Salary" = "double"))
str(base)

readrWay <- read_csv(file = "/Users/deborahchang/Desktop/stat133/stat133-hws-fall17/hw02/data/nba2017-player-statistics.csv", col_names = TRUE, col_types = cols(Player = col_character(), Team = col_character(), Position = col_factor(NULL), Experience = col_character(), Salary = col_double()))
str(readrWay)
```

### 3) Right after importing the data
```{r}
base$Experience <- as.integer(replace(base$Experience, base$Experience == "R", "0"))
```

### 4) Performance of Players
```{r}
base$Missed_FG <- base$FGA - base$FGM
base$Missed_FT <- base$FTA - base$FTM
base$PTS <- 3*base$Points3 + 2*base$Points2 + base$FTM
base$REB <- base$OREB + base$DREB
base$MPG <- base$MIN / base$GP 

base$EFF <- (base$PTS + base$REB + base$AST + base$STL + base$BLK - base$Missed_FG - base$Missed_FT - base$TO) / base$GP
summary(base$EFF)
hist(base$EFF, main = "Histogram of Efficiency (EFF)", xlab = "EFF", ylab = "Frequency")

```

```{r}

sortEFF <- sort(base$EFF, decreasing = TRUE)
topTenByEFFValues <- sortEFF[1:10]
topTenByEFFPlayer <- base$Player[sortEFF[1:10]]
topTenByEFFTeam <- base$Team[sortEFF[1:10]]
topTenByEFFSalary <- base$Salary[sortEFF[1:10]]
frame <- data.frame(topTenByEFFPlayer, topTenByEFFTeam, topTenByEFFSalary, topTenByEFFValues)

frame
```

```{r}
base[base$EFF < 0, 1]
```


```{r}
cor(base$EFF, base$PTS)
cor(base$EFF, base$REB)
cor(base$EFF, base$AST)
cor(base$EFF, base$STL)
cor(base$EFF, base$BLK)
cor(base$EFF, -base$Missed_FG)
cor(base$EFF, -base$Missed_FT)
cor(base$EFF, -base$TO)
```

```{r}
correlations <- c(cor(base$EFF, base$PTS), cor(base$EFF, base$REB),cor(base$EFF, base$AST),cor(base$EFF, base$STL),
+ cor(base$EFF, base$BLK),
+ cor(base$EFF, -base$Missed_FG),
+ cor(base$EFF, -base$Missed_FT),
+ cor(base$EFF, -base$TO))

correlations <- sort(correlations, decreasing = TRUE)
barplot(correlations, main = "Correlations between Player Stats and EFF", width = 0.5, names.arg  = c("PTS", "REB", "STL", "AST", "BLK", "Missed_FT", "Missed_FG", "TO"), las = 2, ylim = c(-1,1), col = c("gray", "gray", "gray", "gray", "gray", "red", "red", "red"))
```

### 5) Efficiency and Salary
```{r}
scatter.smooth(base$EFF, base$Salary, main = "Scatterplot of Efficiency and Salary (in millions of dollars)", xlab = "EFF", ylab = "Salary", col = "blue")
text(20, 1.8*10^7, labels = "lowess")

cor(base$EFF, base$Salary)
```

Based on the correlation between these two variables, there seems to be a somewhat positive relationship between efficiency and salary earned for each player.

```{r}
players2 <- base[base$MPG > 20, ]

scatter.smooth(players2$EFF, players2$Salary, main = "Efficiency and Salary of the more 'Solid' Players", xlab = "EFF", ylab = "Salary", col = "blue")

lines(lowess(players2$EFF, players2$Salary), col = "red")
text(20, 1.8*10^7, labels = "lowess")

cor(players2$EFF, players2$Salary)
```

In the case of the "more established" players, there seems to be less of an association, where other factors aside from those calculated in the EFF formula play a role in performance.

### 6) Comments and Reflections
- Some things that were hard were indexing rows based on certain conditions.
- Some straightforward concepts included creating plots and analyzing these images. 
- I utilized GSI and instructor office hours; the assignment took me about 3 hours.
- The most time consuming part was importing and cleaning the data.
- I still have some questions about row slicing for certain conditions.
- It was exciting to import the data and go through the process of analysis.