---
title: "hw01-deborah-chang.Rmd"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(error = TRUE)
load("data/nba2017-salary-points.RData")
# list available objects
ls()
```


### 1) A bit of data preprocessing
``` {r}
salaryMil <- salary / 10^6
newExp <- as.integer(replace(experience, experience == "R", "0"))
factored <- as.factor(position)
levels(factored) <- c("center", "small_fwd", "power_fwd", "shoot_guard", "point_guard")
table(factored)
```

### 2) Scatterplot of Points and Salary

``` {r}
plot(points, salaryMil, main = "Points vs Salary", xlab = "Points", ylab = "Salary")

```

### 3) Correlation between Points and Salary

``` {r}
numberOfIndividuals <- length(player)
numberOfIndividuals

sumPoints <- sum(points)
sumPoints

sumSalary <- sum(salaryMil)
sumSalary

averagePoints <- sumPoints / length(points)
averagePoints

averageSalary <- sumSalary / length(salaryMil)
averageSalary

variancePoints <- (sum((points - averagePoints)^2)) / (numberOfIndividuals -1)
variancePoints

varianceSalary <- (sum((salary - averageSalary)^2)) / (numberOfIndividuals -1)
varianceSalary

sdevPoints <- sqrt(variancePoints)
sdevPoints

sdevSalary <- sqrt(varianceSalary)
sdevSalary

#covariance
mult <- (points - averagePoints) * (salary - averageSalary)

covariance <- sum(mult) / (numberOfIndividuals -1)
covariance

correlation <- covariance / (sdevPoints*sdevSalary)
corrrelation
```

### 4) Simple Linear Regression

``` {r}
slope <- sum(mult) / sum(points - averagePoints)
slope

intercept <- averageSalary - (slope*averagePoints)
intercept

yHat <- intercept + (slope*points)

summary(yHat)
###write inline equation here
```

The slope coefficient displays the rate at which salary increases by a number of points.
The intercept shows the salary earned when a player scores 0 points. 

- 0 points: 1509886
- 100 points: 2365554
- 500 points: 5788226
- 1000 points: 10066566
- 2000 points: 18623247

### 5) Plotting the regression line 
``` {r}
plot(points, salaryMil, main = "Points vs Salary", xlab = "Points", ylab = "Salary")

```

### 6) Regression residuals and Coefficient of Determination 
### 7) Exploring Position and Experience
### 8) Comments and Reflections
- Some things that were challenging were applying the arguments and functions into my plots.
- Some straightforward items were analyzing the variables through regression and correlation.
- The tutorial was helpful for git and GitHub.
- I utilized office hours and my GSI for guidance in plotting graphs.
- It took over five hours to complete this HW. The most consuming part was troubleshooting.
- I was able to grasp the concepts of this assignment. Most of the assigment was straightforward.
- I was excited to be able to learn how Git and GitHub works and integrates with R Markdown.